---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "tools/README-",
  out.width = "100%"
)
```
# rbin

> Tools for binning data

[![CRAN\_Status\_Badge](http://www.r-pkg.org/badges/version/report)](https://cran.r-project.org/package=rbin) 
[![Travis-CI Build Status](https://travis-ci.org/rsquaredacademy/rbin.svg?branch=master)](https://travis-ci.org/rsquaredacademy/rbin) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/rsquaredacademy/rbin?branch=master&svg=true)](https://ci.appveyor.com/project/aravindhebbali/rbin) [![Coverage status](https://codecov.io/gh/rsquaredacademy/rbin/branch/master/graph/badge.svg)](https://codecov.io/github/rsquaredacademy/rbin?branch=master)  ![](https://img.shields.io/badge/lifecycle-experimental-orange.svg)

## Installation

rbin is not available on CRAN yet. You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("rsquaredacademy/rbin")
```

## Addins

rbin includes two addins for manually binning data:

- `rbinAddin()`
- `rbinFactorAddin()`

## Usage

```{r, echo=FALSE, message=FALSE}
library(rbin)
```

### Manual Binning

```{r manual}
bins <- rbin_manual(mbank, y, age, c(29, 31, 34, 36, 39, 42, 46, 51, 56))
bins 

# plot
plot(bins)
```

### Combine Factor Levels

```{r factor}
# combine levels
upper <- c("secondary", "tertiary")
out <- rbin_factor_combine(mbank, education, upper, "upper")
table(out$education)

# bins
bins <- rbin_factor(out, y, education)
bins 

# plot
plot(bins)
```

### Quantile Binning

```{r quantile}
bins <- rbin_quantiles(mbank, y, age, 10)
bins 

# plot
plot(bins)
```

### Winsorized Binning

```{r winsorize}
bins <- rbin_winsorize(mbank, y, age, 10, winsor_rate = 0.05)
bins 

# plot
plot(bins)
```

### Equal Length Binning

```{r equal_length}
bins <- rbin_equal_length(mbank, y, age, 10)
bins 

# plot
plot(bins)
```

## Alternatives

- [smbinning](https://CRAN.R-project.org/package=smbinning)
- [logiBin](https://CRAN.R-project.org/package=logiBin)
- [woeR](https://CRAN.R-project.org/package=woeR)
- [woeBinning](https://CRAN.R-project.org/package=woeBinning)

## Community Guidelines

Please note that the [34m'rbin'[39m project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.
